### 2018-10-15 MSU-thesis-estee.tex
Split papers from thesis. now just thesis

% This is a template for use with the MSU Thesis class
% Vesion 2.8 2017/12/13
%
% Class options: 
%[PhD]	Doctor of Philosophy (default) 
%[DEd]	Doctor of Education
%[DMA]	Doctor of Musical Arts
%[MA]	Master of Arts
%[MS]	Master of Science
%[MAT]	Master of Arts for Teachers
%[MBA]	Master of Business Administration
%[MFA]	Master of Fine Arts
%[MIPS]	Master of International Planning Studies
%[MHRL]	Master of Human Resources and Labor Relations 
%[MMus]	Master of Music
%[MPH]	Master of Public Health
%[MPP]	Master of Public Policy
%[MSW]	Master of Social Work
%[MURP]	Master in Urban and Regional Planning 
%%
% Default is PhD
%
%
% This template has everything in the right order.
% Just add real content and you're done!
%
\documentclass[]{msu-thesis}
%
% for a prettier, but possibly non-compliant table of contents use the [mixedtoc] option
% for a plain table of contents use the [plaintoc] option
% for a horrendous looking, but possibly required table of contents, use the [boldtoc] option
%
% If you have large tables/figures that need to be in landscape mode, add the [lscape] option

% This is standard fontenc/inputenc for pdflatex
% If you use LuaLaTeX or XeLaTeX you should replace this with the fontspec package
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
%
% If the thesis office requires Times, we'll give them Times
% You can experiment with other font packages here if you like.
% If you are using XeLaTeX or LuaLaTeX load the Times or Times New Roman font with \setmainfont
\usepackage{newtxtext,newtxmath} 
%
% Load any extra packages here
%
% You must specify the title of your thesis, your name, the field of study (not department), and the year
\title{The syntax and semantics of phonology}
\author{Joe Linguist}
\fieldofstudy{Linguistics} % This should be in sentence case
\date{2017}

% If you want a dedication page, specify the text of the dedication here and uncomment the next command.
%
%\dedication{This thesis is dedicated to someone.}
%
\begin{document}

% All the stuff before your actual chapters is called the front matter
\frontmatter
% First make the title page
\maketitlepage
% Next make the abstract
\begin{abstract}
% Your abstract goes here.  Master's 1 page max. PhD 2 page max.
\end{abstract}

% Force a newpage
\clearpage
% Make the copyright page. The Graduate School ridiculously prohibits you
% from having a copyright page unless you pay ProQuest to register the copyright.
% This should be illegal, but I didn't make up the rule.

\makecopyrightpage

% If you have a dedication page, uncomment the next command to print the dedication page
%
%\makededicationpage
%
\clearpage
% Your Acknowledgements are formatted like a chapter, but with no number
\chapter*{Acknowledgements}
\DoubleSpacing % Acknowledgements should be double spaced
Your acknowledgements here.
%
\clearpage
% We need to turn single spacing back on for the contents/figures/tables lists
\SingleSpacing
\tableofcontents* % table of contents will not be listed in the TOC
\clearpage
\listoftables % comment this out if you have no tables
\clearpage
\listoffigures % comment this out if you have no figures
%
% If you have a list of abbreviations/symbols it would go here preceded by a \clearpage
% See the class documentation and the Memoir manual for how to create other lists 
%
\mainmatter
%
% The next line removes the dots in chapter headings in the TOC
% May violate thesis office rules
%\addtocontents{toc}{\protect\renewcommand{\protect\cftchapterdotsep} {\cftnodots}}

\include{chapters/ch_nucleareos}
\include{chapters/ch_piontheory}
\include{chapters/ch_embed}
\include{chapters/ch_theorymodel}
\include{chapters/ch_tpc}
\include{chapters/ch_software}
\include{chapters/ch_experiment}
\include{chapters/ch_analysis}
\include{chapters/ch_results}

%
% If you have pages that must appear in landscape mode, use the [lscape] documentclass option
% and enclose the pages in a {landscape} environment.
%\clearpage\pagestyle{lscape} % first clear the page and change the pagestyle
%\begin{landscape}
%
% your landscape table(s) or figure(s) here
%
%\end{landscape}
%\pagestyle{plain} % remember to change the pagestyle back to plain
%
%
% If you have appendices, they would go here.  
% Comment these lines out if you don't
% If you have more than one appendix, you need to use 
%   \begin{appendices}
%   \chapter{First appendix}
%   \chapter{Second appendix}
%   \end{appendices}
%ch_nucleareos.txt~
\appendix
\chapter{Your appendix}
%
\backmatter
% The next lines add the dots back into the References/Bibliography heading
% of the TOC.  Only uncomment this if you need to put the dots back in having removed them for Chapter headings.
%
%\addtocontents{toc}{%
%   \protect\renewcommand{\protect\cftchapterdotsep} {\cftdotsep}}
%
\makebibliographypage % make the bibliography cover page
% Bibliography can be single spaced
%
\SingleSpacing
%
% Your bibliography command here (e.g. \bibliography{your-bib-file}) if using natbib
%
% Remember that although the bibliography is single spaced, there needs to
% be a blank line between entries. This is set by your bibliography package
% If you are using natbib it is \bibsep; if using biblatex it's \bibitemsep
\end{document}