### 2019-08-07 chapters/ch_intro.tex
Adding better tables and gas section. Macros folder added

\chapter{Introduction}

\section{From Nuclear Forces to the Equation of State}
Figure showing the nuclear forces for the pp, pn, nn 
Figure showing the binding energy vs p-n asym

\section{Phases of Nuclear Matter}

\section{The Nuclear Equation of State}

\section{Boltzmann Ulong Uhlenbeck (BUU) Transport Code}
Figure of transport simulation code showing collision progression

\section{Observables}
Figure showing p-n observable lessens at high energy and from all densities

\section{The $\pi$ observable}
Figure showing delta resonance 
Figure showing pion's produced at 2po 
Figure showing fermi motion effect

\section{Previous Constraints}
Figure showing GW constraint and other previous constraints 
Can you group all the constraints to one nice plot???

\section{Motivation for building the S$\pi$RIT TPC}