### 2020-04-04 chapters/ch_intro.tex

\chapter{Introduction}

\section{From Nuclear Forces to the Equation of State}
Isospin as a good quantum number at low energies 
Figure showing the nuclear forces for the pp, pn, nn 
Pauli exclusion 
inter particle distance , saturation density (maybe figure of all densities?)
Building the infinite matter
Statistical model 
Forces manifest in Energy/particle

\section{The Nuclear Equation of State}
Figure showing the binding energy vs p-n asym
Liquid drop model, mass equation, move to higher densities
Symmetric EoS asymmetric EoS(symmetry energy)
Density dependence of the symmetry energy

\section{Phases of Nuclear Matter}
gas liquid phase, gas, where are we
Progression through the heavy ion collision 
liquid, liquid gas, gas

\section{Studying EoS through Heavy Ion Collisions}
going from infinite matter to finite matter 
approximate nuclear matter with neutron rich radioactive ion beams
probe different asymmetries 
probe different densities with different beam energies 
Symmetry energy goes down at higher beam energies

\section{Boltzmann Ulong Uhlenbeck (BUU) Transport Code}
Does not reach equilibrium for all observable (hadrons) what about pions?
Build up a mean field picture (momentum dependent) unknown quantities here
Non equilibrium can be solved with Boltzman equation with collision term and solved by MC
Figure of transport simulation code showing collision progression 
Clustering is an issue 
Meson, resonance production

\section{Observables of interest to the EoS}
Particle yields of isospin opposites 
Flow??
Figure showing p-n observable lessens at high energy and from all densities

\section{Pion Production }
Figure showing delta resonance 
Figure showing pion's produced at 2po 
Figure showing fermi motion effect (not really sub threshold but ok)
(chemical potential model, delta isobar) prediction for pion ratio
pion mass is small momentum shifted by coulomb affecting shape of pion spectra

\section{Previous Constraints}
Figure showing GW constraint and other previous constraints 
FOPI data at 400 A MeV not as sensitive to symmetry energy 
Saturation density and lower constrained but issues at high density
conflicting analysis on FOPI data which was not intended to be used for Symm Energy
Can you group all the constraints to one nice plot???

\section{Motivation for building the S$\pi$RIT TPC}
Based on EOS TPC
High efficiency for detecting pions and other light charged particles 
Effort to reduce the experimental error bars on such spectra which theory could use