### 2020-04-09 chapters/ch_appendix.tex
Routine Backup

\chapter{Runs analyzed in this data}
\newcommand{\hsn}{$^{132}$Sn+$^{124}$Sn}
\newcommand{\mhsn}{$^{124}$Sn+$^{112}$Sn}
\newcommand{\mlsn}{$^{112}$Sn+$^{124}$Sn}
\newcommand{\lsn}{$^{108}$Sn+$^{112}$Sn}

\begin{table}[H]
  \begin{center}
    \begin{tabular}{ccl}
      \hline
      System & \#runs & Run numbers \\
      \hline\hline
      \hsn & 113 & 2841, 2843, 2844, 2845, 2846, 2848, 2849, 2850, 2851, 2852, 2855, 2856, \\
      & & 2857, 2858, 2859, 2860, 2861, 2875, 2877, 2878, 2879, 2880, 2881, 2882, \\
      & & 2883, 2884, 2887, 2888, 2889, 2890, 2891, 2892, 2893, 2894, 2896, 2898, \\
      & & 2899, 2900, 2901, 2902, 2903, 2904, 2905, 2907, 2914, 2916, 2917, 2919, \\
      & & 2920, 2921, 2922, 2924, 2925, 2926, 2927, 2929, 2930, 2931, 2932, 2933, \\
      & & 2934, 2935, 2936, 2939, 2940, 2941, 2942, 2943, 2944, 2945, 2946, 2948, \\
      & & 2955, 2956, 2958, 2959, 2960, 2961, 2962, 2964, 2965, 2966, 2968, 2969, \\
      & & 2970, 2971, 2972, 2973, 2975, 2976, 2977, 2978, 2979, 2980, 2981, 2982, \\
      & & 2983, 2984, 2985, 2986, 2988, 2989, 2990, 2991, 2992, 2993, 2997, 2999, \\
      & & 3000, 3002, 3003, 3007, 3039 \\
      \hline
      \mhsn & 60 & 2542, 2543, 2544, 2546, 2547, 2548, 2552, 2553, 2554, 2555, 2556, 2557, \\
      & & 2558, 2559, 2560, 2562, 2563, 2564, 2565, 2566, 2567, 2568, 2569, 2570, \\
      & & 2571, 2572, 2573, 2574, 2575, 2578, 2579, 2580, 2581, 2582, 2583, 2584, \\
      & & 2585, 2586, 2587, 2588, 2589, 2590, 2591, 2592, 2593, 2594, 2595, 2596, \\
      & & 2597, 2598, 2599, 2600, 2601, 2617, 2618, 2619, 2620, 2621, 2622, 2623  \\
      \hline
      \mlsn & 68 & 3059, 3061, 3062, 3065, 3066, 3068, 3069, 3071, 3074, 3075, 3076, 3077, \\
      & & 3078, 3080, 3081, 3082, 3083, 3084, 3085, 3087, 3088, 3089, 3090, 3091, \\
      & & 3092, 3093, 3094, 3095, 3097, 3098, 3102, 3103, 3138, 3139, 3140, 3141, \\
      & & 3142, 3143, 3144, 3145, 3146, 3148, 3149, 3150, 3151, 3152, 3153, 3154, \\
      & & 3155, 3156, 3157, 3158, 3159, 3165, 3166, 3167, 3168, 3169, 3170, 3171, \\
      & & 3172, 3177, 3179, 3180, 3181, 3182, 3183, 3184 \\
      \hline
      \lsn & 85 & 2272, 2273, 2274, 2275, 2276, 2283, 2284, 2285, 2286, 2288, 2289, 2291, \\
      & & 2310, 2311, 2314, 2315, 2320, 2322, 2323, 2324, 2325, 2331, 2332, 2333, \\
      & & 2334, 2335, 2336, 2337, 2340, 2341, 2362, 2363, 2368, 2369, 2370, 2371, \\
      & & 2372, 2373, 2374, 2375, 2378, 2379, 2380, 2381, 2382, 2383, 2384, 2385, \\
      & & 2386, 2387, 2388, 2389, 2391, 2392, 2393, 2394, 2395, 2396, 2397, 2398, \\
      & & 2399, 2400, 2401, 2402, 2429, 2432, 2433, 2434, 2437, 2438, 2439, 2440, \\
      & & 2442, 2453, 2461, 2462, 2463, 2501, 2502, 2503, 2505, 2506, 2507, 2508, \\
      & & 2509 \\
      \hline
    \end{tabular}
    \caption{List of runs for the analysis. \label{tb:runList}}
  \end{center}
\end{table}
\chapter{Second appendix}