### 2020-04-28 chapters/ch_data_analysis_2.tex

\chapter{Data Analysis II}

\section{Beam Particle Identification}
Figures of beam contaminants in our beam PID line. 
Table of beam purity reference to Jon's thesis paper

Beam tracking was handled by the BigRIPS spectrometer tracking system. Each individual beam particles coming down the spectrometer was measured using a dE-Time-of-Flight method where the charge, Z, and mass to charge ratio, A/Q, of each particle was measured. Individual particle species represent a two-dimensional Gaussian in this space; both the $\tin{132}{112}$ and $\tin{108}{124}$ systems are shown in Fig.~\ref{fig:beampid}. Clear separation is seen for each particle species and the correct value of Z and A/Q gives us the correct species type. Each species of interest is fitted with multivariate normal distribution with two variables defined as,

\begin{equation}
  f(x,y)=\frac1{2\pi\sigma_x\sigma_y\sqrt{1-\rho^2}}\exp\left\{
  \frac{-(x - \mu_{x})^2/\sigma_x^2-(y-\mu_{y})^2/\sigma_y^2+2\rho
  xy/\sigma_x\sigma_y}{2(1-\rho^2)}\right\},
   \label{multiGauss}
\end{equation}
where x=A/Q, y=Z, $\mu$ is the mean values, and $\sigma$ are the Gaussian widths.

The resulting fit parameters are summarized in Table~\ref{beamParameters}. 
\begin{table}[H]
  \begin{center}
    \begin{tabular}{cccccc}
      \hline 
      Particle Type & $\mu_\mathrm{A/Q}$ & $\sigma_\mathrm{A/Q}$ & $\mu_\mathrm{Z}$ &
      $\sigma_\mathrm{Z}$ & $\rho$\\
      \hline\hline 
      ${}^{132}$Sn & 2.64 & 0.0014 & 49.95 & 0.209 & -0.052 \\
    %  $\tin{124}{112}$ & 2.24005 & 0.00150934 & 49.9906 & 0.194804 & -0.0671999 \\
    %  $\tin{112}{124}$ & 2.47995 & 0.00182327 & 49.961 & 0.214745 & -0.0742297 \\
      ${}^{108}$Sn & 2.16 & 0.0015 & 49.99 & 0.207 & -0.059 \\
      \hline
    \end{tabular}
    \caption{Multivariate normal distribution fit parameters for four beams.
    \label{beamParameters}}
  \end{center}
\end{table}

For both ${}^{132}$Sn an ${}^{108}$Sn a 2.83$\sigma$ cut is taken around the mean values; as seen in Fig.~\ref{beampid132,beampid108}.

\begin{figure}[!htb]
     \centering
     \begin{subfigure}[b]{0.49\textwidth}
         \centering
         \includegraphics[width=\textwidth]{beam-Sn132.png}
         \caption{$\tin{132}{124}$ Beam particle identification.}
         \label{fig:beampid132}
     \end{subfigure}
     \hfill
     \begin{subfigure}[b]{0.49\textwidth}
         \centering
         \includegraphics[width=\textwidth]{beam-Sn108.png}
         \caption{$\tin{132}{124}$ $\pi^-$ angular spectra.}
         \label{fig:beampid108}
     \end{subfigure}
        \caption{$\tin{108}{112}$ Beam particle identification. }
        \label{fig:beampid}
\end{figure}

\section{Edge Cuts}
As seen in Fig.~\ref{fig:edge}, edges of the measurement volume of the TPC we can see several edge effects distortion the clusters positions. The last pads on the edges of the pad-plane naturally have no neighbors on the outsides, therefore their cluster position is biased towards the inside of the TPC. Also since we measure a finite window in time for the vertical direction, pulses near the edge of the time-bucket window are clipped and the reconstructed times are biased towards inside the TPC volume.

The number of affected clusters is small, but the deviation at the end of the track is enough to start causing issues in the momentum reconstruction. Simple cuts around the left,right, top, and bottom of the TPC remove these clusters from the analysis,

\begin{equation*}
  |x|\geq420~\mathrm{mm},\quad y\leq-522+\mathrm{y_o}~\mathrm{mm},
  \quad\mathrm{and}\quad y\geq-64+\mathrm{(Hit\ Shift)}~\mathrm{mm}.
\label{eq:hitshift}
\end{equation*}

\begin{figure}[!htb]
    \centering
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{clusterLR.png} 
        \caption{Generic} \label{fig:clusterLR}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{clusterTop.png} 
        \caption{Competitors} \label{fig:clusterTop}
    \end{subfigure}

    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{clusterBottom.png} 
        \caption{Generic} \label{fig:clusterBottom}
    \end{subfigure}

\label{fig:edge}
\end{figure}

\section{High Density Cut}
The density of tracks near the target region is very high and most of electronics information from this region is inaccurate in both timing and charge value. Hits lying within an semi-ellipsoidal cut region around the target are removed from the software and not included in the track and momentum reconstruction.

Add picture of ellipsoid

\section{Beam angle selection}
The beam is deflected by the magnetic field and impinges on the target at some angle. The BDCs track the beam as a straight line right up until the magnet.  To calculate the incidence angle, the beam is then propagated through the magnetic field until the target position. The beam angle on target is categorized by two angles $\theta_{a_{proj}}$ and $\theta_{b_{proj}}$ defined as,

%define angles 
%define selection

\begin{equation}
  \theta_\mathrm{a,proj}=\tan^{-1}\frac{p_x}{p_z},\quad
  \theta_\mathrm{b,proj}=\tan^{-1}\frac{p_y}{p_z},
  \label{beamAngle}
\end{equation}

where $p_x$, $p_y$, and $p_z$ are the components of beam momentum vector.

Figure~\ref{beamAngleFig} shows the beam angle distribution for ${}^{132}$Sn and ${}^{108}$Sn.

\begin{table}[H]
  \begin{center}
    \begin{tabular}{ccccc}
      \hline 
      System & $\mu_{\theta_\mathrm{a,proj}}$ &
      $\sigma_{\theta_\mathrm{a,proj}}$ & $\mu_{\theta_\mathrm{b,proj}}$ &
      $\sigma_{\theta_\mathrm{b,proj}}$ \\
      \hline\hline 
      $\tin{132}{124}$ & 0.61 & 2.94 & -44.18 & 1.96 \\
    %  $\tin{112}{124}$ & -0.41 & 1.58 & -53.52 & 0.90 \\
      $\tin{108}{112}$ & -0.42 & 1.95 & -55.17 & 0.97 \\
      \hline
    \end{tabular}
    \caption{2D Gaussian fit parameters for $^{132}$Sn, $^{112}$Sn and
      $^{108}$Sn beam angle. \label{beamAngleParameters}}
  \end{center}
\end{table}

\begin{figure}[!htb]
    \centering
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{beamAngle-Sn108.png} 
        \caption{Generic} \label{fig:beamangle108}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{beamAngle-Sn112.png} 
        \caption{Competitors} \label{fig:beamangle112}
    \end{subfigure}

    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{beamAngle-Sn124.png} 
        \caption{Generic} \label{fig:beamangle124}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{beamAngle-Sn132.png} 
        \caption{Competitors} \label{fig:beamangle132}
    \end{subfigure}
\label{fig:beamangle}
\end{figure}

\section{Vertex Cut}
A vertex cut on verticies originating from the target region ensure we are selecting on nuclear collisions with the target. When looking at the z-component of the verticies of all events, we can see several features which correspond to different elements within the beam line and TPC. We also note that near the expected position of the target, WHAT POSITION??, we see a large enhancement of counts which corresponds to nuclear collisions on the target. The z-component of all runs in each beam type are plotted in the range of the target region in Fig.~\ref{fig:vertexdist}. Since the target thickness was less than 1 mm, the vertex resolution achieved can be experimentally measured by fitting the vertex distribution, assuming a Gaussian distribution. The extracted vertex resolutions of each system are summarized in Table~\ref{tb:vertexresol}, with an average vertex resolution of \SI{1.2}{\centi\metre}.

\begin{table*}\centering
\ra{1.3}
\begin{tabular}{@{}rrr@{}}\toprule
\multicolumn{3}{c}{Vertex Resolution}\\
\cmidrule{1-3}
System & Mean (cm) & Sigma (cm) \\
\midrule
$\tin{132}{124}$ & -14.79  & 1.2 \\
$\tin{124}{112}$ & -14.71  & 1.1 \\
$\tin{112}{124}$ & -14.78  & 1.2 \\
$\tin{108}{112}$ & -14.75  & 1.3 \\ 
\bottomrule
\end{tabular}
\caption{Summary of measured verticies and their resolution.}
\label{tb:vertexresol}
\end{table*}

\begin{figure}[!htb]
    \centering
    \begin{subfigure}[t]{\textwidth}
        \centering
        \includegraphics[width=\linewidth]{vertex-Sn108.png} 
        \caption{$\tin{108}{124}$ vertex distributon for all runs.} \label{fig:vertex108}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{\textwidth}
        \centering
        \includegraphics[width=\linewidth]{vertex-Sn112.png} 
        \caption{$\tin{112}{124}$ vertex distributon for all runs.} \label{fig:vertex112}
    \end{subfigure}

    \begin{subfigure}[t]{\textwidth}
        \centering
        \includegraphics[width=\linewidth]{vertex-Sn124.png} 
        \caption{$\tin{124}{112}$ vertex distributon for all runs.} \label{fig:vertex124}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{\textwidth}
        \centering
        \includegraphics[width=\linewidth]{vertex-Sn132.png} 
        \caption{$\tin{132}{124}$ vertex distributon for all runs.} \label{fig:vertex132}
    \end{subfigure}
\label{fig:vertexdist}
\end{figure}

\section{Impact Parameter Selection}
The impact parameter which is defined as the distance (center-to-center), between two nuclei in a collision is a pure theoretical concept. Though experimentally there is no way in experiment to measure the impact parameter directly, several quantities have been indirectly related to the centrality of the collision CITE HERE. As discussed in Section CITE HERE, the Kyoto multiplicity array was used to experimentally trigger on central nuclear collision events. This is under the reasonable assumption that the number of charged particles in a collision grows as the impact parameter decreases. This is best described by a spectator-participant model of nuclear collisions, where a large fraction of the participating nucleons in the overlap region of the two colliding nuclei fragment into individual and clusters of nucleons. In the case where the impact parameter is zero, all the nucleons in both nuclei participate, were as larger impact parameters (peripheral collisions), less nucleons participate.

Certainly the Kyoto array introduces a bias in our measurement of the multiplicity of all the events. At low track multiplicities, the orientation of the event starts to play a major role. Theoretically the orientation of the event (the reaction plane) is random, but due to the fact that we only measure the multiplicity on the sides of the TPC, we are preferentially triggering on events with reaction planes that emit particles in these directions. Therefore events with low track multiplicity (therefore more peripheral), that emit preferentially up and down in the TPC will not be triggered on. This becomes less of an issue as even in mid-peripheral collisions were the number of particles emitted becomes much greater than the Kyoto multiplicity selection.

Assuming a geometric cross section $\sigma$ and the impact parameter of the collision $b$,

\begin{equation}
\sigma = \pi \cdot b^2,
\label{eq:crossSect}
\end{equation}

we can estimate the maximum impact parameter $b_{\mathrm{max}}$, by measuring the total cross section of each system. A detailed analysis on determining the cross section is given in CITE HERE. I have summarized the finding of the cross section and $b_{\mathrm{max}}$ for all the systems. Assuming $b_{\mathrm{max}}$ represents the most peripheral collisions we can define an estimate for the impact parameter $\hat{b}$, which we can define from the cumulative integral of the measured multiplicity distribution,

\begin{equation}
\hat{b} = b_{\mathrm{max}} \frac{\int_{\infty}^{M_x} f(M) dM}{\int_{\infty}^{0} f(M) dM}.
\end{equation}

\section{Track Quality Cuts}
%graphic showing distance to vertex, number of clusters
%number of cluster cut
%distance to vertex
In this section we will discuss the quality of the reconstructed track and will simply refer to it as the ``track quality". Certainly tracks with more clusters are of higher quality than those with few clusters. Occasionally the software will associate clusters and construct a track whose vertex does not coincide with the event vertex. This can occur for a variety of reasons. For a low momentum pion the radius of curvature is such that several rotations can occur in the event which the software can interpret as several tracks; while only one of the reconstructed track appears to originate from the event vertex. It could also happen that false tracks are created by the software from clusters which originate from several different real tracks. These are fairly obvious as they usually traverse the detector horizontally and the number of clusters is very low. In this situation the false track does not originate from anywhere near the vertex.

To reduce the number of fake tracks, or multiple counting of tracks, a simple cut on the measurement of a track's distance to vertex can be used. The vector representing the minimum distance to vertex for each track can be calculated and its corresponding magnitude. We define the distance to vertex cut as a small sphere centered on the vertex location with a radius R. If the distance to vertex of a particular track $d_i$ < R the track is assumed to have originated from the vertex; tracks that are outside are thrown out of the analysis.

Assuming tracks are mostly continuous in clusters, i.e. only randomly missing a few clusters, the number of clusters is directly related to the momentum resolution of a track.

\begin{figure}[!htb]
\includegraphics[width=\textwidth]{distancetovertex.pdf}
\label{fig:poca}
\caption{Distance to vertex for tracks.}
\end{figure}

\section{Angular Quality Cuts}
Since the TPC has no spherical symmetry, there are geometric constraints on measuring tracks. Most notably tracks going up and down in the TPC where the path of the track would be shortest. Tracks traveling up and down in the TPC will have the shortest track length and will correspond to low regions of efficiency. The relevant geometries for this discussion are the corners of the field cage. Figure~\ref{fig:angleEffExplanation} shows a cartoon drawing of the field cage were the angles of each corner is given. Since the window to the field cage was slightly higher, the upper angles are slightly smaller than the lower angles. The $\phi$ angles location of each corner is are $\ang{0}\substack{+24 \\ -34}$ and $\ang{180}\substack{-24 \\ +34}$, with a width of about \ang{58}. While it depends on polar angle $\theta$, a clear cut off around these angles can be seen in Fig.~\ref{fig:angleEffExplanation} in the $\phi$ angles.

The efficiency in these regions are either zero or very small. Typically it is not wise to efficiency correct already small numbers with small efficiency values. Therefore we have taken the practice to cut away the regions of very low efficiencies. The cuts we have made are for $\theta_{C.M.}$ < 90 and for \ang{326} < $\phi$ < \ang{24} and \ang{1} < $\phi$ < \ang{24}.

Maybe add a table of cuts for each system.

We need to correct for the solid angles not measured, or cut away, to arrive at the correct total yields in $4\pi$. The correction is simple given,

\begin{equation}
\Delta\Omega = \Delta\phi(\cos(\theta_1) - \cos(\theta_2)),
\end{equation}

where $\Delta\phi$ is the $\phi$ and $\theta_{1,2}$ are the $\theta$ angular coverage after cuts. The multiplicitive correction factor to the data, C, for the missing angular coverage is simply,

\begin{equation}
C = \frac{4\pi}{\Delta\phi}.
\end{equation}

%As the spherical angle of the track, $\theta$, increases the track gets even shorter and shorter until there are very few clusters at all. We see around $\phi$=\ang{90} and $\phi$=\ang{270}, representing up and down in the TPC respectively, there is near zero efficiency for nearly all angles $\theta$>\ang{30} and for a wide range of $\phi$ angles around these two values.

\begin{figure}[!htb]
\includegraphics[width=\textwidth]{tpcAngle.pdf}
\caption{Rough Geometry of the TPC field cage.}
\label{fig:angleEffExplanation}
\end{figure}

\begin{figure}[!htb]
     \centering
     \begin{subfigure}[b]{0.49\textwidth}
         \centering
         \includegraphics[width=\textwidth]{angularCut-pim-Sn108.png}
         \caption{$y=x$}
         \label{fig:pim108angle}
     \end{subfigure}
     \hfill
     \begin{subfigure}[b]{0.49\textwidth}
         \centering
         \includegraphics[width=\textwidth]{angularCut-pim-Sn132.png}
         \caption{$\tin{132}{124}$ $\pi^-$ angular spectra.}
         \label{fig:pim132angle}
     \end{subfigure}
        \caption{Three simple graphs}
        \label{fig:pim}
\end{figure}

\begin{figure}[!htb]
     \centering
     \begin{subfigure}[b]{0.49\textwidth}
         \centering
         \includegraphics[width=\textwidth]{angularCut-pip-Sn108.png}
         \caption{$\tin{108}{112}$ $\pi^-$ angular spectra.}
         \label{fig:pip108angle}
     \end{subfigure}
     \hfill
     \begin{subfigure}[b]{0.49\textwidth}
         \centering
         \includegraphics[width=\textwidth]{angularCut-pip-Sn132.png}
         \caption{$\tin{132}{124}$ $\pi^-$ angular spectra.}
         \label{fig:pip132angle}
     \end{subfigure}
        \caption{Three simple graphs}
        \label{fig:pip}
\end{figure}