### 2020-06-10 chapters/ch_appendix.tex
Backup

\subsection{Runs analyzed in this data}

\newcommand{\hsn}{$^{132}$Sn+$^{124}$Sn}
\newcommand{\mhsn}{$^{124}$Sn+$^{112}$Sn}
\newcommand{\mlsn}{$^{112}$Sn+$^{124}$Sn}
\newcommand{\lsn}{$^{108}$Sn+$^{112}$Sn}

\begin{table}[!htb]
  \begin{center}
    \begin{tabular}{ccl}
      \hline
      System & \#runs & Run numbers \\
      \hline\hline
      \hsn & 113 & 2841, 2843, 2844, 2845, 2846, 2848, 2849, 2850, 2851, 2852, 2855, 2856, \\
      & & 2857, 2858, 2859, 2860, 2861, 2875, 2877, 2878, 2879, 2880, 2881, 2882, \\
      & & 2883, 2884, 2887, 2888, 2889, 2890, 2891, 2892, 2893, 2894, 2896, 2898, \\
      & & 2899, 2900, 2901, 2902, 2903, 2904, 2905, 2907, 2914, 2916, 2917, 2919, \\
      & & 2920, 2921, 2922, 2924, 2925, 2926, 2927, 2929, 2930, 2931, 2932, 2933, \\
      & & 2934, 2935, 2936, 2939, 2940, 2941, 2942, 2943, 2944, 2945, 2946, 2948, \\
      & & 2955, 2956, 2958, 2959, 2960, 2961, 2962, 2964, 2965, 2966, 2968, 2969, \\
      & & 2970, 2971, 2972, 2973, 2975, 2976, 2977, 2978, 2979, 2980, 2981, 2982, \\
      & & 2983, 2984, 2985, 2986, 2988, 2989, 2990, 2991, 2992, 2993, 2997, 2999, \\
      & & 3000, 3002, 3003, 3007, 3039 \\
      \hline
      \mhsn & 60 & 2542, 2543, 2544, 2546, 2547, 2548, 2552, 2553, 2554, 2555, 2556, 2557, \\
      & & 2558, 2559, 2560, 2562, 2563, 2564, 2565, 2566, 2567, 2568, 2569, 2570, \\
      & & 2571, 2572, 2573, 2574, 2575, 2578, 2579, 2580, 2581, 2582, 2583, 2584, \\
      & & 2585, 2586, 2587, 2588, 2589, 2590, 2591, 2592, 2593, 2594, 2595, 2596, \\
      & & 2597, 2598, 2599, 2600, 2601, 2617, 2618, 2619, 2620, 2621, 2622, 2623  \\
      \hline
      \mlsn & 68 & 3059, 3061, 3062, 3065, 3066, 3068, 3069, 3071, 3074, 3075, 3076, 3077, \\
      & & 3078, 3080, 3081, 3082, 3083, 3084, 3085, 3087, 3088, 3089, 3090, 3091, \\
      & & 3092, 3093, 3094, 3095, 3097, 3098, 3102, 3103, 3138, 3139, 3140, 3141, \\
      & & 3142, 3143, 3144, 3145, 3146, 3148, 3149, 3150, 3151, 3152, 3153, 3154, \\
      & & 3155, 3156, 3157, 3158, 3159, 3165, 3166, 3167, 3168, 3169, 3170, 3171, \\
      & & 3172, 3177, 3179, 3180, 3181, 3182, 3183, 3184 \\
      \hline
      \lsn & 85 & 2272, 2273, 2274, 2275, 2276, 2283, 2284, 2285, 2286, 2288, 2289, 2291, \\
      & & 2310, 2311, 2314, 2315, 2320, 2322, 2323, 2324, 2325, 2331, 2332, 2333, \\
      & & 2334, 2335, 2336, 2337, 2340, 2341, 2362, 2363, 2368, 2369, 2370, 2371, \\
      & & 2372, 2373, 2374, 2375, 2378, 2379, 2380, 2381, 2382, 2383, 2384, 2385, \\
      & & 2386, 2387, 2388, 2389, 2391, 2392, 2393, 2394, 2395, 2396, 2397, 2398, \\
      & & 2399, 2400, 2401, 2402, 2429, 2432, 2433, 2434, 2437, 2438, 2439, 2440, \\
      & & 2442, 2453, 2461, 2462, 2463, 2501, 2502, 2503, 2505, 2506, 2507, 2508, \\
      & & 2509 \\
      \hline
    \end{tabular}
    \caption{List of runs for the analysis. \label{tb:runList}}
  \end{center}
\end{table}

\clearpage

\subsection{Pion Yield Theory}

\begin{table*}[!htb]
\centering
\ra{1.1}
\begin{tabular}{@{}ccccccccccc@{}}
\toprule
        &  & \multicolumn{3}{c}{(a) $^{132}$Sn+$^{124}$Sn} &  & \multicolumn{3}{c}{(b) $^{108}$Sn+$^{112}$Sn} &  \\
        \cmidrule{3-5} \cmidrule{7-9} 
        Code name & $\mathrm{L~(MeV)}$ &  $\mathrm{Y}(\pi^-)$ & $\mathrm{Y}(\pi^+)$ & $\mathrm{SR}(\pi^-/\pi^+)$ &  & $\mathrm{Y}(\pi^-)$  & $\mathrm{Y}(\pi^+)$ & $\mathrm{SR}(\pi^-/\pi^+)$ &  $\mathrm{DR}(\pi^-/\pi^+)$ \\
    \midrule
        $\chi$BUU& 45.6 & 0.509 & 0.109 & 4.67 & & 0.269 & 0.134 & 2.01 & 2.33 \\
        & 120 & 0.483 & 0.117 & 4.13 &  & 0.271 & 0.140 & 1.94 & 2.13 \\
        \addlinespace[0.2cm]
        TuQMD & 54.6 & 0.779 & 0.145 & 5.37 &  & 0.442 & 0.176 & 2.51 & 2.14 \\
        & 145 & 0.839 & 0.145 & 5.79  &  & 0.474 & 0.181 & 2.62 & 2.21 \\
        \addlinespace[0.2cm]
        pBUU & 56.1 & 0.698 & 0.181 & 3.86 &  & 0.401 & 0.213 & 1.88 & 2.05 \\
        & 135 & 0.649 & 0.185 & 3.51  &   & 0.392 & 0.214 & 1.83 & 1.92 \\
        \addlinespace[0.2cm]
        AMD+JAM& 55 & 0.339 & 0.0978 & 3.47 &  & 0.200 & 0.116 & 1.72 & 2.02 \\
        & 152 & 0.311 & 0.0986 & 3.15 &  & 0.192 & 0.116 & 1.66 & 1.90 \\
        \addlinespace[0.2cm]
        IQMD-BNU & 54.6 & 0.542 & 0.148 & 3.67 &  & 0.319 & 0.175 & 1.82 & 2.01 \\
        & 145 & 0.452 & 0.153 & 2.95 &  & 0.278 & 0.167 & 1.67 & 1.77 \\
        \addlinespace[0.2cm]
        SMASH  & 55 & 0.468 & 0.168 & 2.79  &  & 0.287 & 0.190 & 1.51 & 1.85 \\
        & 152 & 0.479 & 0.163 & 2.93 &  & 0.292 & 0.188 & 1.55 & 1.89 \\
        \addlinespace[0.2cm]
        UrQMD & 46 & 0.479 & 0.129 & 3.71 &  & 0.292 & 0.144 & 2.03 & 1.83 \\
        & 104 & 0.449 & 0.133 & 3.38 &  & 0.274 & 0.147 & 1.86 & 1.81 \\
        \addlinespace[0.2cm]
        \bottomrule
    \end{tabular}
    \caption{Pion multiplicities, $Y(\pi^{\pm})$, single ratios $SR(\pi^-/\pi^+)$, and double multiplicity ratios, $DR(\pi^-/\pi^+)$ from seven transport codes. Each code uses two different symmetry energy functions, with all other parameters identical in the codes.}
    \label{tab:pionyieldTheory}
\end{table*}

\clearpage

\chapter{Second appendix}