### 2020-06-19 chapters/ch_appendix.tex
Backup

\subsection{$\Delta$ Resonance Production and Decay Channels}
\label{appen:deltadecay}

The Clebsh-Gordon coefficients for nucleon-nucleon leading to certain $\Delta$ resonances and the respective nucleon: 
\begin{equation}
\begin{split}
& p + p \rightarrow \sqrt{\frac{3}{4}}(\Delta^{++} + n) - \sqrt{\frac{1}{4}}(\Delta^{+} + p) \\
& n + p \rightarrow \sqrt{\frac{1}{2}}(\Delta^{+} + n) - \sqrt{\frac{1}{2}}(\Delta^{0} + p)  \\
& n + n \rightarrow \sqrt{\frac{1}{4}}(\Delta^{0} + n) - \sqrt{\frac{3}{4}}(\Delta^{-} + p) 
\end{split}
\label{eq:deltaProduction}
\end{equation}

A corresponding Clebsh-Gordon coefficients for the decay of $\Delta$ pions into the constituent $\pi$ and corresponding nucleon:

\begin{equation}
\begin{split}
& p + p \rightarrow \sqrt{\frac{5}{6}}(\pi^+ + p + n) - \sqrt{\frac{1}{6}}(\pi^0 + p + p) \\
& n + p \rightarrow \sqrt{\frac{1}{6}}(\pi^+ + n + n) + \sqrt{\frac{2}{3}}(\pi^0 + n + p) + \sqrt{\frac{1}{6}}(\pi^- + p + p)  \\
& n + n \rightarrow \sqrt{\frac{1}{6}}(\pi^0 + n + n) - \sqrt{\frac{5}{6}}(\pi^- + n + p) 
\end{split}
\label{eq:deltaProduction}
\end{equation}

Here we can see that that proton-proton collisions are connected with $\pi^+$ and neutron-neutron collisions are connected with $\pi^-$.

\subsection{Runs analyzed in this data}

\newcommand{\hsn}{$^{132}$Sn+$^{124}$Sn}
\newcommand{\mhsn}{$^{124}$Sn+$^{112}$Sn}
\newcommand{\mlsn}{$^{112}$Sn+$^{124}$Sn}
\newcommand{\lsn}{$^{108}$Sn+$^{112}$Sn}

\begin{table}[!htb]
  \begin{center}
    \begin{tabular}{ccl}
      \hline
      System & \#runs & Run numbers \\
      \hline\hline
      \hsn & 113 & 2841, 2843, 2844, 2845, 2846, 2848, 2849, 2850, 2851, 2852, 2855, 2856, \\
      & & 2857, 2858, 2859, 2860, 2861, 2875, 2877, 2878, 2879, 2880, 2881, 2882, \\
      & & 2883, 2884, 2887, 2888, 2889, 2890, 2891, 2892, 2893, 2894, 2896, 2898, \\
      & & 2899, 2900, 2901, 2902, 2903, 2904, 2905, 2907, 2914, 2916, 2917, 2919, \\
      & & 2920, 2921, 2922, 2924, 2925, 2926, 2927, 2929, 2930, 2931, 2932, 2933, \\
      & & 2934, 2935, 2936, 2939, 2940, 2941, 2942, 2943, 2944, 2945, 2946, 2948, \\
      & & 2955, 2956, 2958, 2959, 2960, 2961, 2962, 2964, 2965, 2966, 2968, 2969, \\
      & & 2970, 2971, 2972, 2973, 2975, 2976, 2977, 2978, 2979, 2980, 2981, 2982, \\
      & & 2983, 2984, 2985, 2986, 2988, 2989, 2990, 2991, 2992, 2993, 2997, 2999, \\
      & & 3000, 3002, 3003, 3007, 3039 \\
      \hline
      \mhsn & 60 & 2542, 2543, 2544, 2546, 2547, 2548, 2552, 2553, 2554, 2555, 2556, 2557, \\
      & & 2558, 2559, 2560, 2562, 2563, 2564, 2565, 2566, 2567, 2568, 2569, 2570, \\
      & & 2571, 2572, 2573, 2574, 2575, 2578, 2579, 2580, 2581, 2582, 2583, 2584, \\
      & & 2585, 2586, 2587, 2588, 2589, 2590, 2591, 2592, 2593, 2594, 2595, 2596, \\
      & & 2597, 2598, 2599, 2600, 2601, 2617, 2618, 2619, 2620, 2621, 2622, 2623  \\
      \hline
      \mlsn & 68 & 3059, 3061, 3062, 3065, 3066, 3068, 3069, 3071, 3074, 3075, 3076, 3077, \\
      & & 3078, 3080, 3081, 3082, 3083, 3084, 3085, 3087, 3088, 3089, 3090, 3091, \\
      & & 3092, 3093, 3094, 3095, 3097, 3098, 3102, 3103, 3138, 3139, 3140, 3141, \\
      & & 3142, 3143, 3144, 3145, 3146, 3148, 3149, 3150, 3151, 3152, 3153, 3154, \\
      & & 3155, 3156, 3157, 3158, 3159, 3165, 3166, 3167, 3168, 3169, 3170, 3171, \\
      & & 3172, 3177, 3179, 3180, 3181, 3182, 3183, 3184 \\
      \hline
      \lsn & 85 & 2272, 2273, 2274, 2275, 2276, 2283, 2284, 2285, 2286, 2288, 2289, 2291, \\
      & & 2310, 2311, 2314, 2315, 2320, 2322, 2323, 2324, 2325, 2331, 2332, 2333, \\
      & & 2334, 2335, 2336, 2337, 2340, 2341, 2362, 2363, 2368, 2369, 2370, 2371, \\
      & & 2372, 2373, 2374, 2375, 2378, 2379, 2380, 2381, 2382, 2383, 2384, 2385, \\
      & & 2386, 2387, 2388, 2389, 2391, 2392, 2393, 2394, 2395, 2396, 2397, 2398, \\
      & & 2399, 2400, 2401, 2402, 2429, 2432, 2433, 2434, 2437, 2438, 2439, 2440, \\
      & & 2442, 2453, 2461, 2462, 2463, 2501, 2502, 2503, 2505, 2506, 2507, 2508, \\
      & & 2509 \\
      \hline
    \end{tabular}
    \caption{List of runs for the analysis. \label{tb:runList}}
  \end{center}
\end{table}

\clearpage

\subsection{Pion Yield Theory}
\label{tb:pionyieldTheory}

\begin{table*}[!htb]
\centering
\ra{1.1}
\begin{tabular}{@{}ccccccccccc@{}}
\toprule
        &  & \multicolumn{3}{c}{(a) $^{132}$Sn+$^{124}$Sn} &  & \multicolumn{3}{c}{(b) $^{108}$Sn+$^{112}$Sn} &  \\
        \cmidrule{3-5} \cmidrule{7-9} 
        Code name & $\mathrm{L~(MeV)}$ &  $\mathrm{Y}(\pi^-)$ & $\mathrm{Y}(\pi^+)$ & $\mathrm{SR}(\pi^-/\pi^+)$ &  & $\mathrm{Y}(\pi^-)$  & $\mathrm{Y}(\pi^+)$ & $\mathrm{SR}(\pi^-/\pi^+)$ &  $\mathrm{DR}(\pi^-/\pi^+)$ \\
    \midrule
        $\chi$BUU& 45.6 & 0.509 & 0.109 & 4.67 & & 0.269 & 0.134 & 2.01 & 2.33 \\
        & 120 & 0.483 & 0.117 & 4.13 &  & 0.271 & 0.140 & 1.94 & 2.13 \\
        \addlinespace[0.2cm]
        TuQMD & 54.6 & 0.779 & 0.145 & 5.37 &  & 0.442 & 0.176 & 2.51 & 2.14 \\
        & 145 & 0.839 & 0.145 & 5.79  &  & 0.474 & 0.181 & 2.62 & 2.21 \\
        \addlinespace[0.2cm]
        pBUU & 56.1 & 0.698 & 0.181 & 3.86 &  & 0.401 & 0.213 & 1.88 & 2.05 \\
        & 135 & 0.649 & 0.185 & 3.51  &   & 0.392 & 0.214 & 1.83 & 1.92 \\
        \addlinespace[0.2cm]
        AMD+JAM& 55 & 0.339 & 0.0978 & 3.47 &  & 0.200 & 0.116 & 1.72 & 2.02 \\
        & 152 & 0.311 & 0.0986 & 3.15 &  & 0.192 & 0.116 & 1.66 & 1.90 \\
        \addlinespace[0.2cm]
        IQMD-BNU & 54.6 & 0.542 & 0.148 & 3.67 &  & 0.319 & 0.175 & 1.82 & 2.01 \\
        & 145 & 0.452 & 0.153 & 2.95 &  & 0.278 & 0.167 & 1.67 & 1.77 \\
        \addlinespace[0.2cm]
        SMASH  & 55 & 0.468 & 0.168 & 2.79  &  & 0.287 & 0.190 & 1.51 & 1.85 \\
        & 152 & 0.479 & 0.163 & 2.93 &  & 0.292 & 0.188 & 1.55 & 1.89 \\
        \addlinespace[0.2cm]
        UrQMD & 46 & 0.479 & 0.129 & 3.71 &  & 0.292 & 0.144 & 2.03 & 1.83 \\
        & 104 & 0.449 & 0.133 & 3.38 &  & 0.274 & 0.147 & 1.86 & 1.81 \\
        \addlinespace[0.2cm]
        \bottomrule
    \end{tabular}
    \caption{Pion multiplicities, $Y(\pi^{\pm})$, single ratios $SR(\pi^-/\pi^+)$, and double multiplicity ratios, $DR(\pi^-/\pi^+)$ from seven transport codes. Each code uses two different symmetry energy functions, with all other parameters identical in the codes.}
    \label{tab:pionyieldTheory}
\end{table*}

\clearpage

\section{Dalitz Decay of the $\pi^0$}
\label{appen:dalitz}

\section{Cut Variation Analysis}
\label{sec:cutvar}
The particular track quality cuts described in Section~\ref{sec:qualitycut} are there to reduced the contamination from poorly reconstructed tracks which contribute to the background in the PID spectra. The best set of values were found for all the data sets which include the charged particle multiplicity of the event, dOCA, and the number of clusters cut. By varying the cuts in both the data and the MC efficiency, we can evaluate the analysis to see if there are any systematic dependencies on the observable we plot. But varying the cuts undoutably means as the cut get tighter, and less data is included, the statistical error bar will increase making it difficult to compare to more loose cuts. This is mitigated by looking at the uncorrelated error described in \cite{dataAnalysis}. Here if the systematic trend of the observable is much larger than the statistical error bars on the default cuts and the uncorrelated error bars, then there is some systematic trend that either exists due to physics or some error or miscalculation in the analysis. It is usually not recommended to estimated systematic error bars using this method, but sometimes is the only way.

Here we will discuss the particular analysis using the total pion ratio as the example. Figure~\ref{fig:totalRatioError} shows the total pion as a function of several cut variations, where one cut is varied and all the others are held constant. Three cuts are of particular interest, the event multiplicity, dOCA, and the number of cluster cut. Figure~\ref{fig:clustTotalRatio} shows the variation as the number of cluster cut is varied. The default cut used is > 20 clusters which is represented by the middle point. The red bars represent the statistical error of this default value. The error bars on the other points are the uncorrelated error bars as described by the prescription in \cite{dataAnalysis}. Here it is clear that there is a systematic bias in the pion ratio as a function

%show the pim total yield 
%discuss others in the apendix

\begin{figure}[!htb]
     \centering
     \begin{subfigure}[b]{\textwidth}
         \centering
         \includegraphics[width=\textwidth]{132Sn_Multiplicity_NoEC_pimTotal}
         \caption{No efficiency correction.}
         \label{fig:multCutVar}
     \end{subfigure}
     \hfill
    \begin{subfigure}[b]{\textwidth}
         \centering
         \includegraphics[width=\textwidth]{132Sn_Multiplicity_pimTotal}
         \caption{Efficiency corrected.}
         \label{fig:multCutVar}
     \end{subfigure}
     \hfill
\caption{Y($\pi^-$) when varying the ${}^{132}$Sn charged particle multiplicity cut. }
\label{fig:cutVar}
\end{figure}

\begin{figure}[!htb]
     \centering
     \begin{subfigure}[b]{\textwidth}
         \centering
         \includegraphics[width=\textwidth]{Clusters_NoEC_pimTotal}
         \caption{Before efficiency correction.}
         \label{fig:pim_clustVar_NoEC}
     \end{subfigure}
     \hfill
    \begin{subfigure}[b]{\textwidth}
         \centering
         \includegraphics[width=\textwidth]{Clusters_pimTotal}
         \caption{After efficiency correction.}
         \label{fig:pim_clustVar_EC}
     \end{subfigure}
     \hfill
\caption{Y($\pi^-$) when varying the number of cluster cut of the tracks.}
\label{fig:pim_clustVar}
\end{figure}

\begin{figure}[!htb]
     \centering
     \begin{subfigure}[b]{\textwidth}
         \centering
         \includegraphics[width=\textwidth]{POCA_NoEC_pimTotal}
         \caption{No efficiency correction.}
         \label{fig:pim_cutVarPOCA_NoEC}
     \end{subfigure}
     \hfill
    \begin{subfigure}[b]{\textwidth}
         \centering
         \includegraphics[width=\textwidth]{POCA_pimTotal}
         \caption{Efficiency corrected.}
         \label{fig:pim_cutVarPOCA_EC}
     \end{subfigure}
     \hfill
\caption{Y($\pi^-$) when varying the dOCA cut.}
\label{fig:pim_cutVarPOCA}
\end{figure}

\begin{figure}[!htb]
     \centering
     \begin{subfigure}[b]{\textwidth}
         \centering
         \includegraphics[width=\textwidth]{132Sn_Multiplicity_NoEC_pipTotal}
         \caption{No efficiency correction.}
         \label{fig:pip_multCutVar}
     \end{subfigure}
     \hfill
    \begin{subfigure}[b]{\textwidth}
         \centering
         \includegraphics[width=\textwidth]{132Sn_Multiplicity_pipTotal}
         \caption{Efficiency corrected.}
         \label{fig:pip_multCutVar}
     \end{subfigure}
     \hfill
\caption{Y($\pi^+$) when varying the ${}^{132}$Sn charged particle multiplicity cut. }
\label{fig:cutVar}
\end{figure}

\begin{figure}[!htb]
     \centering
     \begin{subfigure}[b]{\textwidth}
         \centering
         \includegraphics[width=\textwidth]{Clusters_NoEC_pipTotal}
         \caption{Before efficiency correction.}
         \label{fig:pip_clustVar_NoEC}
     \end{subfigure}
     \hfill
    \begin{subfigure}[b]{\textwidth}
         \centering
         \includegraphics[width=\textwidth]{Clusters_pipTotal}
         \caption{After efficiency correction.}
         \label{fig:pip_clustVar_EC}
     \end{subfigure}
     \hfill
\caption{Y($\pi^+$) when varying the number of cluster cut of the tracks.}
\label{fig:pim_clustVar}
\end{figure}

\begin{figure}[!htb]
     \centering
     \begin{subfigure}[b]{\textwidth}
         \centering
         \includegraphics[width=\textwidth]{POCA_NoEC_pipTotal}
         \caption{No efficiency correction.}
         \label{fig:pip_cutVarPOCA_NoEC}
     \end{subfigure}
     \hfill
    \begin{subfigure}[b]{\textwidth}
         \centering
         \includegraphics[width=\textwidth]{POCA_pipTotal}
         \caption{Efficiency corrected.}
         \label{fig:pip_cutVarPOCA_EC}
     \end{subfigure}
     \hfill
\caption{Y($\pi^+$) when varying the dOCA cut.}
\label{fig:pip_cutVarPOCA}
\end{figure}

\begin{figure}[!htb]
     \centering
     \begin{subfigure}[b]{\textwidth}
         \centering
         \includegraphics[width=\textwidth]{132Sn_Multiplicity_NoEC_TotalRatio}
         \caption{No efficiency correction.}
         \label{fig:ratio_multCutVar}
     \end{subfigure}
     \hfill
    \begin{subfigure}[b]{\textwidth}
         \centering
         \includegraphics[width=\textwidth]{132Sn_Multiplicity_TotalRatio}
         \caption{Efficiency corrected.}
         \label{fig:ratio_multCutVar}
     \end{subfigure}
     \hfill
\caption{Y($\pi^+$)/Y($\pi^-$) when varying the ${}^{132}$Sn charged particle multiplicity cut. }
\label{fig:cutVar}
\end{figure}

\begin{figure}[!htb]
     \centering
     \begin{subfigure}[b]{\textwidth}
         \centering
         \includegraphics[width=\textwidth]{Clusters_NoEC_TotalRatio}
         \caption{Before efficiency correction.}
         \label{fig:ratio_clustVar_NoEC}
     \end{subfigure}
     \hfill
    \begin{subfigure}[b]{\textwidth}
         \centering
         \includegraphics[width=\textwidth]{Clusters_TotalRatio}
         \caption{After efficiency correction.}
         \label{fig:ratio_clustVar_EC}
     \end{subfigure}
     \hfill
\caption{Y($\pi^+$)/Y($\pi^-$) when varying the number of cluster cut of the tracks.}
\label{fig:pim_clustVar}
\end{figure}

\begin{figure}[!htb]
     \centering
     \begin{subfigure}[b]{\textwidth}
         \centering
         \includegraphics[width=\textwidth]{POCA_NoEC_TotalRatio}
         \caption{No efficiency correction.}
         \label{fig:ratio_cutVarPOCA_NoEC}
     \end{subfigure}
     \hfill
    \begin{subfigure}[b]{\textwidth}
         \centering
         \includegraphics[width=\textwidth]{POCA_TotalRatio}
         \caption{Efficiency corrected.}
         \label{fig:ratio_cutVarPOCA_EC}
     \end{subfigure}
     \hfill
\caption{Y($\pi^+$)/Y($\pi^-$) when varying the dOCA cut.}
\label{fig:pip_cutVarPOCA}
\end{figure}

\chapter{Second appendix}