### 2020-07-16 MSU-thesis-estee.tex
Almost final draft. Before Bill's edits for Intro, Results, abstract, summary.

% This is a template for use with the MSU Thesis class
% Vesion 2.8 2017/12/13
%
% Class options: 
%[PhD]	Doctor of Philosophy (default) 
%[DEd]	Doctor of Education
%[DMA]	Doctor of Musical Arts
%[MA]	Master of Arts
%[MS]	Master of Science
%[MAT]	Master of Arts for Teachers
%[MBA]	Master of Business Administration
%[MFA]	Master of Fine Arts
%[MIPS]	Master of International Planning Studies
%[MHRL]	Master of Human Resources and Labor Relations 
%[MMus]	Master of Music
%[MPH]	Master of Public Health
%[MPP]	Master of Public Policy
%[MSW]	Master of Social Work
%[MURP]	Master in Urban and Regional Planning 
%%
% Default is PhD
%
%
% This template has everything in the right order.
% Just add real content and you're done!
%
\documentclass[]{msu-thesis}

%
% for a prettier, but possibly non-compliant table of contents use the [mixedtoc] option
% for a plain table of contents use the [plaintoc] option
% for a horrendous looking, but possibly required table of contents, use the [boldtoc] option
%
% If you have large tables/figures that need to be in landscape mode, add the [lscape] option

%IS THIS THE RIGHT BIB STYLE TO USE????!?!?!??!?!?!??!
%!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
\bibliographystyle{elsarticle-num}

% This is standard fontenc/inputenc for pdflatex
% If you use LuaLaTeX or XeLaTeX you should replace this with the fontspec package
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{float}
\usepackage{graphicx}
\usepackage{cleveref}
\usepackage{overpic}
\usepackage{contour}
\usepackage{color}
\usepackage{amsmath}
\usepackage{cleveref}
\usepackage{rotating}
\usepackage{mathtools}
\usepackage{appendix}
\usepackage{comment}\renewcommand{\arraystretch}{1.5}
\usepackage[version=3]{mhchem}

\graphicspath{{images/spacecharge/} {images/} {images/mccomparison/}  {images/dataAnalysis_2/} {images/dynamicrange/} {images/dataAnalysis/} {images/experiment/} {images/results/}  {images/intro/} }

%
% If the thesis office requires Times, we'll give them Times
% You can experiment with other font packages here if you like.
% If you are using XeLaTeX or LuaLaTeX load the Times or Times New Roman font with \setmainfont
\usepackage{newtxtext,newtxmath,booktabs,siunitx,caption,subcaption}

%\sisetup{separate-uncertainty}
\sisetup{input-uncertainty-signs}

%
% Load any extra packages here
%
% You must specify the title of your thesis, your name, the field of study (not department), and the year
\title{Constraining the High Density Nuclear Symmetry Energy with Pions}
\author{Justin Brian Estee}
\fieldofstudy{Physics} % This should be in sentence case
\date{2020}

% If you want a dedication page, specify the text of the dedication here and uncomment the next command.
%
\dedication{}
%
%User defined commands
\newcommand{\spirit}{S$\pi$RIT }
\newcommand{\numerr}[3]{ \num[parse-numbers=false]{#1 \pm #2 (sys.) \pm #3 (stat.)}  }
%\newcommand{\numerr}[3]{ \num[input-close-uncertainty]{#1 \pm #2 (sys.) \pm #3 (stat.)}  }

\newcommand{\tin}[2]{{}^{#1}\mathrm{Sn} + {}^{#2}\mathrm{Sn}}
\newcommand{\ra}[1]{\renewcommand{\arraystretch}{#1}}

\makeatletter
\DeclareTextCommand{\textprime}{\encodingdefault}{%
  \mbox{$\m@th'\kern-\scriptspace$}% 
}
\makeatother

\DeclareSIUnit\torr{Torr}
\DeclareSIUnit\eVperc{\eV\per\clight}
\DeclareSIUnit\MeVA{A \mega\eV }
\DeclareSIUnit\clight{\text{\ensuremath{c}}}
\DeclareSIUnit\micron{\micro\metre}
\DeclareSIUnit\mrad{\milli\rad}
\DeclareSIUnit\gauss{G}

\includeonly{
chapters/ch_intro,
%ch_theory,
chapters/ch_experiment,
chapters/ch_data_analysis,
chapters/ch_data_analysis_2,
chapters/ch_results,
chapters/ch_summary,
chapters/ch_appendix
}

\begin{document}
%\sisetup{quotient-mode=fraction} % Output a/b as \frac{a}{b}

% All the stuff before your actual chapters is called the front matter
\frontmatter
% First make the title page
\maketitlepage
% Next make the abstract
\begin{abstract}
% Your abstract goes here.  Master's 1 page max. PhD 2 page max.

Recent astronomical measurements of neutron star mergers have reinvigorated questions about the nature of dense matter in the universe. The densities reached in these stars range from  normal nuclear densities $\rho_o$ to upwards of $9\rho_o$, where $\rho_o = \SI{1.7e14}{\gram\per\centi\metre\squared}$, extremely dense matter, which is primarily composed of neutrons. The symmetry energy describes the difference in binding energy between pure neutron matter and symmetric matter with equal number of neutrons and protons. The density dependence of the symmetry energy is related to the pressure, which supports the star preventing it from collapsing under its own gravitation force. For years, theory and experiment  have combined to progressively constrain the equation of state of symmetric matter, and it was in the last decade that the density dependence of pure neutron matter (the symmetry energy), had been constrained at normal nuclear densities and lower, $\rho < \rho_o$. Still large uncertainties still remain at densities around 2$\rho_o$ which are important to the understanding the properties of neutron stars.

Besides measuring neutrons stars directly, the laboratory provides the only way to study dense nuclear matter at different densities, and to probe different proton-neutron asymmetries. Charged pions are produced from nucleon-nucleon collisions through the short-lived intermediate $\Delta$ resonance, in the early, dense parts of heavy pion collisions. Yet, existing experimental data sensitive to the symmetry energy was lacking, prompting the construction of the \spirit Time Projection Chamber (TPC) which was constructed to measure charged pion spectra ($\pi^\pm$) with a high efficiency produced in neutron-rich heavy-ion collisions. This dissertation encompasses the first pion spectra from the most neutron-rich ($\tin{132}{124}$), and neutron-poor ($\tin{108}{112}$) systems. Also we will compare the results with the most recent transport  simulations. We will highlight some of the successes and hope to motivate a discussion in the theoretical community on how to better reproduce the pion phenomena in neutron-rich heavy-ion collisions, which is important for not only constraining the density dependence of the symmetry energy, but also for understanding the $\Delta$ and pion roles in neutron stars.

\end{abstract}

% Force a newpage
\clearpage
% Make the copyright page. The Graduate School ridiculously prohibits you
% from having a copyright page unless you pay ProQuest to register the copyright.
% This should be illegal, but I didn't make up the rule.

\makecopyrightpage

% If you have a dedication page, uncomment the next command to print the dedication page
%
%\makededicationpage
%
\clearpage
% Your Acknowledgements are formatted like a chapter, but with no number
\chapter*{Acknowledgements}
\DoubleSpacing % Acknowledgements should be double spaced
Your acknowledgements here.
%
\clearpage
% We need to turn single spacing back on for the contents/figures/tables lists
\SingleSpacing
\tableofcontents* % table of contents will not be listed in the TOC
\clearpage
\listoftables % comment this out if you have no tables
\clearpage
\listoffigures % comment this out if you have no figures
%
% If you have a list of abbreviations/symbols it would go here preceded by a \clearpage
% See the class documentation and the Memoir manual for how to create other lists 
%
\mainmatter
%
% The next line removes the dots in chapter headings in the TOC
% May violate thesis office rules
%\addtocontents{toc}{\protect\renewcommand{\protect\cftchapterdotsep} {\cftnodots}}
\chapterstyle{default}

\include{chapters/ch_intro}
%\include{chapters/ch_theory}
\include{chapters/ch_experiment}
\include{chapters/ch_data_analysis}
\include{chapters/ch_data_analysis_2}
\include{chapters/ch_results}
\include{chapters/ch_summary}

%
% If you have pages that must appear in landscape mode, use the [lscape] documentclass option
% and enclose the pages in a {landscape} environment.
%\clearpage\pagestyle{lscape} % first clear the page and change the pagestyle
%\begin{landscape}
%
% your landscape table(s) or figure(s) here
%
%\end{landscape}
%\pagestyle{plain} % remember to change the pagestyle back to plain
%
%
% If you have appendices, they would go here.  
% Comment these lines out if you don't
% If you have more than one appendix, you need to use 
 %  \begin{appendices}
  % \chapter{First appendix}
  % \chapter{Second appendix}
  % \end{appendices}

\appendix
%\addappheadtotoc
\include{chapters/ch_appendix}

%
\backmatter
% The next lines add the dots back into the References/Bibliography heading
% of the TOC.  Only uncomment this if you need to put the dots back in having removed them for Chapter headings.
%
%\addtocontents{toc}{%
%   \protect\renewcommand{\protect\cftchapterdotsep} {\cftdotsep}}
%
\makebibliographypage % make the bibliography cover page
% Bibliography can be single spaced
%
\SingleSpacing
%
% Your bibliography command here (e.g. \bibliography{your-bib-file}) if using natbib
%
% Remember that although the bibliography is single spaced, there needs to
% be a blank line between entries. This is set by your bibliography package
% If you are using natbib it is \bibsep; if using biblatex it's \bibitemsep
\bibliography{mybibfile}

\end{document}